---
title: "Explore_and_Join_PFAS"
author: "Kyle P Messier"
date: "2023-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, packages, echo=FALSE}
library(data.table)
library(ggplot2)
library(tidyverse)
library(magrittr)
library(janitor)
library(sf)
library(cartography)
library(nhdplusTools)


```


## Reading in Data

Here, we read in the PFAS data we obtained from Rebecca Fry at UNC.

1) PFAS Monitoring

We then read in geospatial data:

1) Public water supply sources (polygons)
2) Surface water intakes (points)
3) Current Public Water systems  (polygons)
4) NC Surface Water Supply Watersheds (polygons)
5) EPA Federal Registered Facilities in NC

```{r Read in Data, echo= FALSE}
#### Read in the PFAS Data ####
PFAS <- fread("input/PFAST-Summary-Round1.csv")


#### Read in the GIS Data from NCOneMap ####

##Public Water Supply Sources ##
Public.Water.Supply.Sources <- sf::read_sf("input/GIS/Public_Water_Supply_Water_Sources/Public_Water_Supply_Water_Sources.shp")

## Surface_Water_Intakes ##
Surface.Water.Intakes <- sf::read_sf("input/GIS/Surface_Water_Intakes/Surface_Water_Intakes.shp")

## Current Public Water Systems ## 
Current.Public.Water.Systems <- sf::read_sf("input/GIS/Type_A_Current_Public_Water_Systems_(2004)/Type_A_Current_Public_Water_Systems_(2004).shp")

## Surface Water Supply Watersheds ## 
Surface.Watersheds <- sf::read_sf("input/GIS/NC_Surface_Water_Supply_Watersheds/dwq_wsws_20141126.shp")


## EPA Federal Registered Facilities in NC
FRS <- sf::read_sf("input/GIS/state_single_nc/STATE_SINGLE_NC.csv")

# FRS <- fread("input/GIS/state_single_nc/STATE_SINGLE_NC.csv")  
# FRS.Water <- subset(FRS,SITE_TYPE_NAME == "WATER SYSTEM")

```

## Create coordinates in consistent projected and geographic coordinate systems
 
Transform and/or calculate consistent spatial coordinates
The public water supply sources uses the NC State Plane coordinate system
we will use this as the common projected coordinate system
We will also calculate lat/lon coordinates
 


```{r,echo=FALSE}
 
PWS.crs <- st_crs(Public.Water.Supply.Sources)

PWS.coords.proj <- Public.Water.Supply.Sources %>% st_coordinates() %>% as_tibble()
PWS.coords.geo <- Public.Water.Supply.Sources %>% 
                  st_transform(crs = 4269) %>% 
                  st_coordinates() %>% as_tibble() %>% 
                  rename("Longitude" = "X","Latitude" = "Y")

Public.Water.Supply.Sources <- cbind(Public.Water.Supply.Sources,PWS.coords.proj,PWS.coords.geo)

```

## Fix unmatched ID

There are many mismatches between the EPA ID number and the ID number in the 
PFAS data. We go through by hand and make the necessary changes so 
that we do not lose data/locations when we do the join 


 Prior to Joining we have to fix some wrong IDs in the PFAS data
 StationID and EPA_srcID of the unmatched PFAS data
 We need to correct these to the correct public water source ID (EPA_srcID)
 We'll do this by hand
 
 [1,] "NC0100010-2" "39012"
 [2,] "NC0161015"   "18601"
 [3,] "NC0286020-2" "65263"
 [4,] "NC0363025-2" "39297"
 [5,] "NC0407030"   "31427"
 [6,] "NC0424020"   "32565"
 [7,] "NC0431010"   "33929"
 [8,] "NC0433015"   "34137"
 [9,] "NC0437020"   "34248"
 [10,] "NC0446015"   "34565"
 [11,] "NC0452010"   "34686"
 [12,] "NC0465020"   "35340"
 [13,] "NC0465025"   "35349"
 [14,] "NC0466040"   "35860"
 [15,] "NC6059015-1" "71963"


Here we fix the matches that are relatively straight forward
```{r, Fix Straighforwad Mismatches, echo=FALSE}


 # FRS.Water.coords.proj <- FRS.Water %>% sf::st_coordinates()

# NC0100010-2 --> simple mismatch, change 39012 to 39013
PFAS$EPA_srcID[PFAS$EPA_srcID== 39012] = 39013

#NC0161015  -->  Multiple wells with the IDs  different. 
#                Change it to the one that looks to be the primary source

PFAS$EPA_srcID[PFAS$EPA_srcID== 18601] = 39013

# NC0286020-2 --> NC0286020 is a match to the town. Change the EPA_srcID to match
PFAS$EPA_srcID[PFAS$EPA_srcID== 65263] = 39251

# NC0363025-2 --> simple mismatch in EPA_srcID (also remove the -2 in station ID)
PFAS$EPA_srcID[PFAS$EPA_srcID== 39297] = 39296


# "NC0424020"   "32565" --> --> Multiple wells with the IDs just slightly different. 
#                       Change it to the one that looks to be the primary source
PFAS$EPA_srcID[PFAS$EPA_srcID== 32565] = 32566

# "NC0431010"   "33929" --> Multiple wells with the IDs just slightly different. 
#                       Change it to the one that looks to be the primary source
PFAS$EPA_srcID[PFAS$EPA_srcID== 33929] = 33932

# NC0433015"   "34137"--> Multiple wells with the IDs just slightly different. 
#                       Change it to the one that looks to be the primary source
PFAS$EPA_srcID[PFAS$EPA_srcID== 34137] = 34139

# "NC0437020"   "34248"--> Multiple wells with the IDs just slightly different. 
#                       Change it to the one that looks to be the primary source
PFAS$EPA_srcID[PFAS$EPA_srcID== 34248] = 34250

# "NC0446015"   "34565"--> Multiple wells with the IDs just slightly different. 
#                       Change it to the one that looks to be the primary source
PFAS$EPA_srcID[PFAS$EPA_srcID== 34565] = 34566


# "NC0465020"   "35340"--> Multiple wells with the IDs just slightly different. 
#                       Change it to the one that looks to be the primary source
PFAS$EPA_srcID[PFAS$EPA_srcID== 35340] = 35339


#"NC0465025"   "35349"--> Multiple wells with the IDs just slightly different. 
#                       Change it to the one that looks to be the primary source
PFAS$EPA_srcID[PFAS$EPA_srcID== 35349] = 35350

# "NC0466040"   "35860"--> Multiple wells with the IDs just slightly different. 
#                       Change it to the one that looks to be the primary source
PFAS$EPA_srcID[PFAS$EPA_srcID== 35860] = 35861

# "NC6059015-1" "71963"-->  IDs just slightly different. 
#                       Change it to the one that looks to be the primary source
PFAS$EPA_srcID[PFAS$EPA_srcID== 71963] = 74912


```




## Join with the public water supply wells
```{r, Join, echo=FALSE}

# Do the join 
# Drop unnecessary columns

PFAS$EPA_srcID <- as.numeric(PFAS$EPA_srcID)
PFAS.Water.Join <- left_join(PFAS,Public.Water.Supply.Sources,by = c("EPA_srcID"= "federal_id")) |> 
   select(-c(objectid,pws_id,county,owner,state_id,source_nam,reg_name,system_nam,system_add,system_cit,system_sta,system_zip,
             system_con,system_pop,owner_addr,owner_city,owner_stat,owner_zip,wsw_class,intake_loc,source_ava,well_yield,well_depth,X,Y))



```


## Fix the more troublesome IDs 

```{r fix the troublesome IDs with direct substitution, echo=FALSE}
# NC0407030  31427 --> Can join using the water system name in FRS (BATH WATER SYSTEM) Has lat/lon
#381241
# Public.Water.Supply.Sources[Public.Water.Supply.Sources station_name == "BATH WATER SYSTEM"),]
# REGISTRY_ID = 110027381241
# We directly substitute the long/lat into the PFAS and Water Supply joined data

PFAS.Water.Join[PFAS.Water.Join$station_name=="BATH WATER SYSTEM",]$Longitude = as.numeric(FRS[FRS$REGISTRY_ID==110027381241,]$LONGITUDE83)
PFAS.Water.Join[PFAS.Water.Join$station_name=="BATH WATER SYSTEM",]$Latitude = as.numeric(FRS[FRS$REGISTRY_ID==110027381241,]$LATITUDE83)


# "NC0452010"   "34686"
# This is the town of Maysville - it does not match to the public water supply or FRS
# It is also not mapped on the SDWIS or FRS website. It is however, near Camp Lejune 
# and has many his PFAS, so we need to resolve it 

# Coordinates provided by the Town of Maysville, Town Manager
maysville.lat <- 34.907347
maysville.lon <- -77.232910


PFAS.Water.Join[PFAS.Water.Join$station_id=="NC0452010",]$Longitude = maysville.lon
PFAS.Water.Join[PFAS.Water.Join$station_id=="NC0452010",]$Latitude = maysville.lat

# The PFAS and public water supply intake join

PFAS.Water.Supply <- sf::st_as_sf(PFAS.Water.Join)


```



## Let's join the PFAS data with the water supply distribution area
```{r}

PFAS.Water.Distribution <- sf::st_join(PFAS.Water.Supply,Current.Public.Water.Systems,left=FALSE,join = st_nearest_feature)


# make a plot of the unique distribution areas
PFAS.Dist.Plot <- sf::st_difference(st_geometry(PFAS.Water.Distribution))
plot(PFAS.Dist.Plot)
```





