---
title: "NC Public Water Utilities"
author: "Kyle P Messier"
date: "2023-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, packages, echo=FALSE, include=FALSE}
library(data.table)
library(ggplot2)
library(tidyverse)
library(magrittr)
library(janitor)
library(sf)
library(mapsf)
library(nhdplusTools)


```


## North Carolina Water Utility Data Sources

1) Public Water Source Intakes
These are point locations of well intakes for public water sources. 
I have joined these to our PFAS data using a unique EPA assigned ID

Obtained from NCOneMap at https://arcg.is/iD8vy

2) Type A Current Public Water Systems 
These are polygons that are supposed to represent the entire area of a 
public water system. 

Obtained from NCOneMap at https://arcg.is/1Ge8zm

```{r Read in Data, echo= FALSE}

## Water Supply joined with PFAS Data

PFAS.join <- sf::st_read("input/PFAS_Public_Source_Join.gpkg")

## Current Public Water Systems ## 
Current.Public.Water.Systems <- sf::read_sf("input/GIS/Type_A_Current_Public_Water_Systems_(2004)/Type_A_Current_Public_Water_Systems_(2004).shp")


```


The following map shows the North Carolina Public Water systems with 
the colors representing their ownership types. 


```{r, echo=FALSE}
# set a theme
mf_theme("default")
# plot a shadow
mf_init(Current.Public.Water.Systems)
# plot 

mypal <- mf_get_pal(11, "spectral", rev = TRUE)
mf_map(
  x = Current.Public.Water.Systems,
  var = "waownty",
    type = "typo",
  leg_pos = "bottomleft2",
  leg_title = "Ownership Type",
    pal = mypal,
  border = "grey30",
  add = TRUE
)


# layout
mf_layout(
  title = "NC Public Water Systems",
  credits = paste0(
    "Sources: NC One Map\n",
    "mapsf ",
    packageVersion("mapsf")
  )
)

```

Here, we look at the PWS that are only municipal. These areas look much 
more in line with what I would expect as a water distribution area.

```{r,echo=FALSE}

CPWS.muni <- Current.Public.Water.Systems[Current.Public.Water.Systems$waownty=="Municipality",]

# set a theme
mf_theme("default")
# plot a shadow
# mf_shadow(CPWS.muni)
# plot municipalities
mf_init(CPWS.muni)
# mf_map(PFAS.join, add = TRUE)

mf_map(CPWS.muni, add = TRUE)
# plot population

# mf_map(
#   x = CPWS.muni,
#   var = "waownty",
#     type = "typo",
#   leg_pos = "bottomleft2",
#   leg_title = "Ownership Type",
#     pal = mypal,
#   border = "grey30"  
# )


# layout
mf_layout(
  title = "Municipal Public Water Systems",
  credits = paste0(
    "Sources: NC One Map\n",
    "mapsf ",
    packageVersion("mapsf")
  )
)

```


Questions: 

1) Is it unlikely that an entire county would be supplied by a public water
utility like those in eastern NC that are an entire county?

2) Is it normal to have a well or intake location outside the boundary of the 
supply area? i.e. some of the points are outside the municipal area

3) If you were to estimate the total areas in NC serviced by a public water utility,
would you use just the municipal or also include some of the other ownership types?



